### 2016-08-14 writeup/RochatEtal_ResponseToReviewers.tex
Starts ResponseToReviewers document

\documentclass[11pt]{article}

%% WRY has commented out some unused packages %%
%% If needed, activate these by uncommenting
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\geometry{a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
%\geometry{landscape}                % Activate for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent

\usepackage{natbib}

%for figures
%\usepackage{graphicx}

\usepackage{color}
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
%% for graphics this one is also OK:
\usepackage{epsfig}

%% AMS mathsymbols are enabled with
\usepackage{amssymb,amsmath}

%% more options in enumerate
\usepackage{enumerate}
%\usepackage{enumitem}

%% insert code
\usepackage{listings}

\usepackage[utf8]{inputenc}
\usepackage{multicol}
\setlength{\columnsep}{1cm}

\usepackage{hyperref}

\usepackage{fourier, heuristica}
\usepackage{array, booktabs}
\usepackage{graphicx}
\usepackage[x11names]{xcolor}
\usepackage{colortbl}
\usepackage{caption}
\DeclareCaptionFont{blue}{\color{LightSteelBlue3}}

\newcommand{\foo}{\color{LightSteelBlue3}\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}}

\newcommand{\R}[1]{\textbf{R}\textbf{#1}}
\newcommand{\ie}{i.e.}

\newcommand{\sZ}{\mathsf{Z}}
\newcommand{\sE}{\mathsf{E}}
%\newcommand{\sA}{\mathsf{A}}

\title{{\normalfont Response to reviewers of the manuscript \texttt{2016GL070261} }
       ``Seasonality of submesoscale
       dynamics \\in the Kuroshio Extension''}
\author{Cesar B. Rocha,
        Sarah T. Gille, Teresa K. Chereskin, \& Dimitris Menemenlis}
\date{}
\begin{document}

%\include{symbols}

\maketitle

\section{Reviewer \#1}

{\it This is a timely investigation into the submesoscale sea surface height (SSH) and
velocity variations in the Kuroshio Extension region using global high-resolution
OGCM simulations. The main conclusion of the study that the surface ocean submesoscale
turbulence and inertia-gravity waves modulate out-of-phase seasonally, is novel and
important. Overall, the manuscript is well presented and clearly written and I
recommend its publication in GRL after the following comments are addressed.}\\

We thank reviewer \#1 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\begin{enumerate}

\item {\it In the bottom paragraph on P.10, the authors emphasized that "At scales larger
than 20km, the April and October spectra based on hourly velocity snapshots are
nearly indistinguishable from each other". This emphasis appears to be contradictory
to the statements in other places of the manuscript that the seasonality of subinertial
submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}

\item {\it Does the mesoscale eddy variability in the Kuroshio Extension region of the model
have a similar energy level as those in the along-track altimeter data? This point
is worth mentioning as it impacts on the relative importance between the instability-induced
turbulence versus intia-gravity waves.}

\item {\it P.4, line 2 from bottom: Missing a word after "early".}

\item {\it P.6, line 5-7: Should October 15 be part of "late spring/early summer"?}

\item {\it P.7, line 12: Change "KE spectra" to "KE spectrum"?}

\item {\it P.8, bottom: The black line in Figure 2c shows a non-zero divergence value.
      Shouldn't the gridded SSH data be non-divergent by definition?}

\item {\it P.11, line 4-6: Figure 4a reveals that daily averaging not only reduces the
      kinetic energy level for scales smaller than 250km, but also those longer
      than 500km. What part of the large-scale flow signals are being removed?}

\end{enumerate}

%
% Reviewer 2
%

\section{Reviewer \#2}
{\it In this paper, the authors describe seasonality of submesoscale (10-100 km)
activities in Kuroshio Extension region by using outputs from their new high
resolution numerical simulations. Authors found that submesoscale turbulence
dominates in late winter and early spring because of mixed layer instabilities, and
inertial-internal waves in late summer and early fall because of internal tides and
stratification near the surface. This new simulation, which includes tides and a high
frequency wind forcing, and authors' attempts to analyze surface submesoscale
turbulence and near-inertial waves are potentially important to oceanographic
community. However, some discussions seem weak, possibly because it is very
difficult to separate sub- and super-inertial motions in the submeso dynamics as
authors mentioned.}\\

We thank reviewer \#2 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\subsection{Major points}

\begin{enumerate}

  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
        and early fall is unclear to me. In the paper, authors attribute the mechanism to
        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
        any interaction mechanism of internal tides (please at least add some reference). With
        the fixed frequencies of tidal constituents and existence of the critical latitude for the
        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
        wave generation within the whole analyzed region by internal tides (with the
        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
        estimated from the non-tidal forcing run in the Supplemental material, authors'
        conclusion could be strengthened.}

  \item {\it I also wonder why authors ignore wind-induced near-inertial waves (X-13, the first
        line). It is well studied that wind-induced near-inertial motion in mixed layer in back
        ground flow fields disperse quickly and propagate into the ocean interior (e.g., Young
        and Ben-Jelloul 1997, J. Mar. Res., 55, 735-66). The distribution of those waves
        depends on the distribution of vorticity (or Laplacian of vorticity) of background
        flows (Klein and Llewellyn Smith 2001, J. Mar. Res., 59, 697-723; Klein et al. 2004, Q.
        J. R. Meteorol. Soc., 130, 1153-1166). Therefore, those wind-induced near-inertial
        waves could have the horizontal scales of 10-100km. Note that those studies are based
        on geostrophically balanced or quasigeostrophic turbulent field. However, authors
        may need to study interactions between submeso dynamics regime (Rossby number
        and Richardson number are close to 1) and near-inertial waves in winter and spring
        cases where I guess that non-linear terms may play a role. I don't think that authors
        need to explain all internal wave dynamics in this GRL paper. I would rather suggest
        to remove all discussions related to submesoscale inertial-gravity waves and focus on
        describing differences between October and April.}

  \item {\it Finally, I don't understand why authors define 10-100 km as a submesoscale and used
        this scale for the spatial filtering (X-7, section 3). Authors should discuss Figure 4 first
        and then give a reason why they define 10-100 km as a submesoscale and use this
        scale for the filtering. In X-11, section 5, author point out that a part of the dynamics
        around 250 km scale should be explained by near-inertial waves. Then, I wonder why
        authors decide to discuss internal waves in 10-100 km scale.}
\end{enumerate}

\subsection{Minor points}

\begin{enumerate}

  \item {\it X-4, section 1, 2nd paragraph and X-13, section 6 last: Authors claimed that they
        analyze the Kuroshio Extension region. However, their study region does not
        include between $~$140E and $~$155E where the Kuroshio Extension is strongest in
        Figure 1(a). Please explain why you exclude this region in the paper. Since those
        longitudes includes the Izu-Ogasawara ridge (internal tide generation) and
        formation region of Subtropical mode water (strong mixed layer front), I would
        expect that including these latitudes would strengthen authors conclusions. I
        would also like to point out that positions of KEO and KESS moorings in the
        supplemental material are outside of the study region.}

  \item {\it X-4, section 1, 3rd paragraph: “submesoscale inertial-gravity waves”. I feel that this
        nomenclature is too sudden. It might be better to write “inertial-gravity waves
        with a horizontal scales of 10-100 km (hereafter submesoscale inertial-gravity
        waves)”?}

  \item {\it X-9, section 4, 2nd paragraph: I wonder what we could learn from the sentence “In
          other words, even in April, when submesoscale turbulence prevails, only the
          daily-averaged fields are largely in geostrophic balance”. This suggests that the
          balanced and upward cascade motion (originated from the submesoscale
          instability) scenario in Sasaki et al., (2014, nature comm. 5) could be wrong
          because they use the daily averaged data (X-11, section 5, last of 1st paragraph)?
          Or it just means that the balanced regime is actually the submeso-dynamics
          regime?}

  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}

\end{enumerate}

\section{Reviewer \# 3}
        {\it This manuscript provides model-based evidence of a seasonal cycle in submesoscale motions
         in the western North Pacific (Kuroshio Extension). The manuscript summarizes the results of
         a study in which two years of moderate-resolution (1/24th degree or 4.6 km) and one year of
         high-resolution (1/48th degree or 2.3 km) model output were examined for evidence of
         submesoscale activity. Measures of submesoscale activity include vorticity, divergence and
         strain rate and the authors demonstrate that these quantities are enhanced in winter. Kinetic
         energy and sea surface height (SSH) spectra also show seasonal trends. These results are
         consistent with an existing suite of studies reporting seasonality to submesoscale flows.

         What is relevant for the present study and distinguishes it from previous studies is that the
         authors report an increase in internal wave activity during summer and which (they suggest)
         projects onto the SSH variance. Because of the relevance of this study to challenges facing
         the upcoming Surface Water Ocean Topography (SWOT) mission—which aims at measuring
         submesoscale flows from space from nearly-balanced flows—this study will be of interest to
         readers of GRL. The major error that should be addressed prior to publication is a more direct
         connection of divergence and increased SSH variance to inertia-gravity waves. At present,
         this connection is entirely speculative since any ageostrophic flows can give rise to
        divergence and SSH variance. Edits should be made to the text and figures, as well.}\\

        We thank  Christian Buckingham for the speedy and thorough review. Below we address
                  Buckingham's comments and suggestions.

\subsection{Major comments}

\begin{enumerate}

  \item {\it The authors must clearly demonstrate that increased inertia-gravity waves (IGWs) are
        responsible for the seasonal signal in horizontal divergence and SSH spectra at high
        wavenumbers. The authors argue that this is so but it is not demonstrated with sufficient
        clarity. It is true that a portion of this energy/variance will be attributed to internal waves
        modified by rotation but other flows give rise to divergence (and thus, SSH variance). Thus,
        the seasonal changes reported by the authors might be the result of seasonal changes in these
        other flows. Given that this is the main message of the paper, the authors must make this
        modification.}

  \item {\it There is very little introduction in this manuscript. The authors must address this
          shortcoming prior to publication. The obvious method of addressing this issue is to discuss
          the seasonality of IGWs in the context of SWOT and why this might matter. The introduction
          should, therefore, place this study matter?”}

  \item {\it There needs to be a discussion about the effective resolution of your model. The fact that
        the model has a resolution of dx = 2.3 km and 4.6 km does not mean it can resolve oceanic
        phenomena at this scale. Its effective horizontal resolution is closer to 8 times the grid
        resolution (Soufflet et al. 2013). This corresponds to 18 km and 35 km for LLC4320 and
        LLC2160 output, respectively. [A more exact value could be obtained by estimating zonal
        and meridional spectra (of u, v or SSH) and identifying the high wavenumbers / low
        wavelengths at which the spectral energy falls precipitously.] This means the model is not
        resolving a number of submesoscale coherent vortices (SCVs) that may occur as a result of
        ML baroclinic instability; these have diameters close to 5 km. So the follow-up questions
        from a dynamical point-of-view become (1) where does the unresolved energy come from or
        go to? and (2) how does this affect your results? These questions need to be addressed or at
        least discussed if the main topic of your paper is submesoscale dynamics.}

 \item {\it If the authors could estimate the magnitude of the sea surface height signature within a
        particular band as a function of season that might be helpful. For example, the authors could
        integrate SSH spectra (Figure 4b) from hourly fields between 10 km and 33 km (i.e.,
        wavenumber = 3 x 10-2 km-1 which appears to be the cross-over point for the two curves) to
        obtain SSH variances within a particular wavenumber band. Additionally, under the
        assumption that SSH variance in winter is dominated by balanced motions while SSH
        variance in summer is the sum of balanced + unbalanced motions, the difference between
        these two quantities yields a bound1 for the SSH variance due to unbalanced motions.}

\end{enumerate}

\subsection{Minor comments}

\begin{enumerate}

  \item {\it (Section 1, Paragraph 1) The authors use the term ``submesoscale'' and ``mesoscale'' as
        descriptive of motions having (1-100-km) and (100-300 km) horizontal scales,
        respectively. I find these definitions misleading.

        One way to define mesoscale and submesoscale is to define what scales of motion a
        particular sensor will resolve. While this would not be optimal, it is objective. This is
        probably the motivation for the definition used by Callies and Ferrari (2013), since the
        swath altimeter (as initially proposed) would resolve scales greater than 1 km but
        smaller than the present-day, profiling satellite altimeters – i.e., eddy-like features
        with diameters greater than 80-100 km (see Chelton et al. 2011, Progress).
        Unfortunately, this definition has the adverse effect of changing for each sensor.

        A better definition of the submesoscale might be founded on dynamical arguments.
        Wunsch and Stammer (1998) introduce the geostrophic equations and note that the
        lower limit for which this approximation is strictly valid is ~30 km. This conveniently
        defines the lower limit of the mesoscale regime. I would therefore define the
        submesoscale regime as anything smaller than this but one for which the hydrostatic
        approximation remains valid (e.g., 1 km).

      The motivation for wanting to use the term ``submesoscale'' in the title is clear –
      submesoscale motions are a hot topic within the community and it catches individuals’
      eyes. This is fine. But what must be made clear within the manuscript is what scales of
      motion and what force balances (i.e., dynamical regimes) are being described.}

  \item {\it (Abstract) The abstract could be rewritten to emphasize its connection to SWOT.}

  \item {\it (Section 1, last paragraph) Adding “with implications for SWOT” at the end of the
          introduction would also help this article.}

  \item {\it (Section 1, Paragraph 1) The authors mention a suite of papers documenting
        seasonality at the submesoscale but fail missed a few: Ostrovskii (1995) and
        Brannigan et al. (2015). These are observation- and model-based studies, resp.}

  \item {\it (Section 1, Paragraph 2) Model resolutions: it would be helpful for readers to place in
        parentheses nominal horizontal resolutions corresponding to these spatial resolutions.}

  \item {\it (Section 1, Paragraph 3) What types of horizontal and temporal scales for IGWs are
        we talking about here? How would this fall onto the Garrett-Munk (GM) internal
        wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
        flows, not just internal waves, despite that the name suggests this.)}

 \item {\it Discretization of the underlying equations of motion. How do you expect the
        combination of vertical and horizontal resolution modifies internal wave properties
        that exist in your model? Might you be attributing characteristics to these waves that
        would be different in the real ocean?}

\item {\it There should be some discussion of the seasonality of the mixed layer deformation
      radius. Buckingham et al. (2016) point out that all modeling and observational studies
      to date necessarily introduce seasonally varying energy by virtue of the deformation
      radius changing with season and falling below the grid resolution in certain seasons.
      This might be worth a comment.}

\item {\it (Section 2, Paragraph 1) ``The LLC4320 simulation is an extension of the 3-month
      long output used by Rocha et al. (2016).'' I think this is good to mention, here, so
      please keep. In contrast, in (Section 5, Paragraph 2) the authors note a consistency
      between the results in the Kuroshio Extension and those in the Drake Passage. This is
      not a consistency as it is a different location, different time. The author has already
      made the reader aware of the previous study so there is no reason to cite again.}

\item {\it (Section 5, Paragraph 1) eliminate ``spectral'' after Hanning window. Also, no need for
        quotation marks.}

\item {\it (Section 5, Paragraph 1) ``the projection of these flows onto different horizontal scales
            ...'' Have you thought about computing scale-dependent vorticity, strain rate and
            divergence? This is really what you would be doing
            aforementioned flows into spectral space.}

\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}

\item {\it (Section 4, first paragraph) I understand the motivation for examining the Laplacian of
      sea surface height but the average reader will not understand the connection.}

\item {\it (Section 4, second paragraph) Last sentence. This is good. I understand this but to
        make this legible for the general audience it should be explained why we expect a near
        one-to-one line for the joint-PDF of vorticity and the Laplacian of SSH.}

\item {\it (Section 4, last paragraph) ``consistent with linear inertia-gravity waves''.
        It would be
        helpful to have a reference after this statement.}

\item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}

\item {\it (Section 4, last paragraph) ``...whereas divergence is moderately, negatively skewed as
        predicted by ...'' Specifically mention convergence/downwelling.}

\item {\it (Figure 1) ** Place panels (d) and (e) before panels (b) and (c) since this is the order
        in which you refer to them in the manuscript. ** Difficult to see contours in (d) and
        (e). ** Consider overlaying mixed layer depth on these transects since it is mentioned
        in the text.}

\item {\it (Figure 2) Is the mean of the strain rate zero? Is the mean of the divergence zero? The
        variance should be used rather than the root-mean-squared (RMS).}

\item {\it (Figure 2) What is the sensitivity of the estimates to the stencil type? See Arbic et al.
        paper describing this sensitivity for satellite altimetry.}

\item {\it (Section 3 and Figures 2 \& 3) At what depth are you computing these quantities?}

\item {\it (Section 2, last paragraph) It is not just solar radiation that enters the surface buoyancy
      flux calculation.}

\item {\it (Section 2, last paragraph) ``as shallow as 40 m'' – I bet that the mixed layer depths get
      even shallower than this. At midlatitudes, 20-30 m is not uncommon.}

\item {\it 24. (Section 3, Paragraph 1) ``These diagnostic highlight the submesoscale structures in
      the flow.'' It might help the authors to motivate their calculations using the following:

      Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
      strain rate and mean flow. This can be seen by expressing uh as a Taylor series
      expansion and then decomposing the velocity gradient into symmetric and anti-
      symmetric parts [Landau and Lifshitz, 1987]. Because vortices, vertical motion and
      fronts are ubiquitous at the submesoscale, this is a natural decomposition and
      variances of these quantities are apt descriptors of submesoscale turbulence.}

\item {\it (Figure 3) Would be nice to include a third row that contains the collapsed (i.e., one-
      dimensional) PDFs of normalized vorticity to illustrate the skewness of hourly and
      daily-averaged flow. I believe you already have this information in the supplementary
      material but would be nice to illustrate to the reader. Also, in your normalization of
      vorticity, are you accounting for the fact that the Coriolis parameter changes with
      latitude? Your domain spans considerable changes in f and the relevant dynamical
      parameter is the gradient Rossby number.}

\item {\it (Figure 4b) Strictly speaking, the label on the y-axis should not be SSH variance. SSH
      variance would be the integral of this curve. The y-axis should be labelled as the
      power spectral density of sea surface height (SSH), or SSH Spectra.}

\item {\it The authors may find the work of Brüggeman and Eden (2015) helpful in addressing
      some of the aforementioned questions.}

\end{enumerate}

\end{document}