### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Response to minor comments.

\documentclass[11pt]{article}

%% WRY has commented out some unused packages %%
%% If needed, activate these by uncommenting
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\geometry{a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
%\geometry{landscape}                % Activate for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent

\usepackage{natbib}
\bibliographystyle{chicago}

\usepackage{color}

%for figures
%\usepackage{graphicx}

\usepackage{color}
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
%% for graphics this one is also OK:
\usepackage{epsfig}

%% AMS mathsymbols are enabled with
\usepackage{amssymb,amsmath}

%% more options in enumerate
\usepackage{enumerate}
%\usepackage{enumitem}

%% insert code
\usepackage{listings}

\usepackage[utf8]{inputenc}
\usepackage{multicol}
\setlength{\columnsep}{1cm}

\usepackage{hyperref}

\usepackage{fourier, heuristica}
\usepackage{array, booktabs}
\usepackage{graphicx}
\usepackage[x11names]{xcolor}
\usepackage{colortbl}
\usepackage{caption}
\DeclareCaptionFont{blue}{\color{LightSteelBlue3}}

\newcommand{\foo}{\color{LightSteelBlue3}\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}}

\newcommand{\R}[1]{\textbf{R}\textbf{#1}}
\newcommand{\ie}{i.e.}

\newcommand{\sZ}{\mathsf{Z}}
\newcommand{\sE}{\mathsf{E}}
%\newcommand{\sA}{\mathsf{A}}

\newcommand{\bdp}{\begin{description}}
\newcommand{\edp}{\end{description}}

\title{{\normalfont Reply to reviewers of the manuscript \texttt{2016GL070261} }
       ``Seasonality of submesoscale
       dynamics \\in the Kuroshio Extension''}
\author{Cesar B. Rocha,
        Sarah T. Gille, Teresa K. Chereskin, \& Dimitris Menemenlis}
\date{}
\begin{document}

%\include{symbols}

\maketitle

\section*{Summary}

\section{Reviewer \#1}

{\it The authors have adequately addressed all the questions raised in my
previous review. I recommend the manuscript be accepted for publication in GRL.}

\bdp
  \item We thank the reviewer for constructive review.
\edp

\section{Reviewer \#2}

{\it I have read the revised manuscript and the authors have adequately addressed
     my concerns. Here I would like to point out things I noticed.
     Please ignore those if you don't like.}

\bdp
   \item We thank the reviewer for constructive review.
\edp

\begin{enumerate}

\item {\it Line 166. "super-inertial IGWs" I think IGW represents any oscillation whose
                frequency is between f and N. If so, super-inertial is not needed.}

\item {\it Line 174. "inertia gravity waves" should be IGWs.}

\item {\it Line 187. I am not sure if PSI is important near surface and wonder if it
                is resolved in this model because of the vertical resolution.
                Authors could discuss about PSI in the model with Jenn Mackinnon there.}

\item {\it Line 196. Remove "?"}

\item {\it In supplement, page x-3 first line. Remove "r"}

\item {\it To separate IGWs and other submesoscale motion, Ertel vorticity could be
      used (e.g. Kunze 1993, JPO, 2567-2588). This paper is about deep ocean though.}

\end{enumerate}

\section{Reviewer \#3}

\bibliography{rocha_etal.bib}

\end{document}